Method

### 2025-11-26 GroupPolarizationStatmod/paper/method.qmd
old updates to method

To evaluate an experimental design's ability to properly identify null effects,
we need to simulate participant responses assuming the mere agreement null model
that invalidated the replication, if one was found. We simulate data from a
single group polarization experiment by sampling participants' ordinal opinions
at pre-/post-discussion from the pre- and post-discussion opinion distributions
specified by the null model that invalidated the replication in that experiment,
$\nu_e$. If there are $N_e$ participants in experiment $e$ then our simulated
data sets contained $2N_e$ rows. We recorded $N_e$ in the web app as metadata
along with data about average opinions and opinion shifts for each experimental
condition across the corpus.

It can be helpful to

@@

Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is
then inspected to see if spurious group polarization occurred in the simulation.
Spurious group polarization is determined by first inferring the latent
parameters of each simulated experimental trial by fitting an ordered probit
model [@Liddell2018]. The procedure infers the latent distribution parameters
that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$,
$\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.

The inferred parameters are then used to calculate the effect size for that
trial, Cohen's $d$, defined as\~\cite[p. 331]{Liddell2018} $$
  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
$$ {#eq-cohens} It is implied that these values are the inferred values, not the
values from $\nu$, when used in the context of calculating Cohen's $d$. Note
$d=0$ always for the true $\nu$.

>>>>

To evaluate an experimental design's ability to properly identify null effects,
we need to simulate participant responses assuming the mere agreement null model
that invalidated the replication, if one was found. We simulate data from a
single group polarization experiment by sampling participants' ordinal opinions
at pre-/post-discussion from the pre- and post-discussion opinion distributions
specified by the null model that invalidated the replication in that experiment,
$\nu_e$. If there are $N_e$ participants in experiment $e$ then our simulated
data sets contained $2N_e$ rows: two time step rows per participant. We recorded
$N_e$ used by each experiment in the web app as metadata along with data about
average opinions and opinion shifts for each experimental condition across the
corpus.

Group polarization is "detected" if the group polarization effect size meets or
exceeds a threshold value. We use Cohen's $d$ for the effect size, defined as
[@Liddell2018, 331] $$
  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
$$ It is implied that these values are the inferred values, not the values from
$\nu$, when used in the context of calculating Cohen's $d$. Note $d=0$ always
for the true $\nu$.

@@
